---
title: "DDSAnalytics Salary and Attrition Analysis"
author: "Laura Lazarescou"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

# Exploratory Data Analysis (EDA)
## Analyze variables for single values or lack of relevancy

We removed the following variables from the original dataset:
  Over18
  StandardHours 
  EmployeeCount 
  EmployeeNumber
  
## Convert numeric classifications to factors (for R compatibility)


```{r}



#
# R Code for DDS Project
#

library(tidyverse)
library(GGally)
library(fastDummies)
library(dplyr)
library(mlbench)
library(caret)
library(mlr)
library(ggthemes)
library(gplots)
library(randomForest)
library(skimr)
library(corrplot)
library(e1071) # Naive Bayes Classifier
# Many thanks for Naive Bayes and Visualization code from Zulaikha Lateef on https://www.edureka.co/blog/naive-bayes-in-r/

# Read data sets (3) and copy for manipulation

CS <- read.csv("C:/SMU MS Data Science Files/SMU Downloads/6306 DDS/Project 2 DDSAnalytics/CaseStudy2-data.csv")
CS2 <- CS

TestNoAttrition <- read.csv("C:/SMU MS Data Science Files/SMU Downloads/6306 DDS/Project 2 DDSAnalytics/CaseStudy2CompSet No Attrition.csv")
TestAtt <- TestNoAttrition

TestNoSalary <- read.csv("C:/SMU MS Data Science Files/SMU Downloads/6306 DDS/Project 2 DDSAnalytics/CaseStudy2CompSet No Salary2.csv")
TestSal <- TestNoSalary

# Exploratory Data Analysis (EDA) 

str(CS2)
summary(CS2)

# Remove variables where all the same value
# Over18=Y, StandardHours=80, EmployeeCount, EmployeeNumber not relevant to salary or attrition

CS2$Over18 <- NULL
CS2$StandardHours <- NULL
CS2$EmployeeCount <- NULL
CS2$EmployeeNumber <- NULL

# Check NA and missing values
sum(is.na(CS2))
sum(is.na(TestNoAttrition))
sum(is.na(TestNoSalary))

# No NAs in reference data sets

# Convert categorical data to factors

CS2$JobInvolvement <- as.factor(CS2$JobInvolvement)
CS2$JobLevel <- as.factor(CS2$JobLevel)
CS2$JobSatisfaction <- as.factor(CS2$JobSatisfaction)
CS2$PerformanceRating <- as.factor(CS2$PerformanceRating)
CS2$RelationshipSatisfaction <- as.factor(CS2$RelationshipSatisfaction)
CS2$StockOptionLevel <- as.factor(CS2$StockOptionLevel)
CS2$TrainingTimesLastYear <- as.factor(CS2$TrainingTimesLastYear)
CS2$WorkLifeBalance <- as.factor(CS2$WorkLifeBalance)
CS2$Education <- as.factor(CS2$Education)
CS2$EnvironmentSatisfaction <- as.factor(CS2$EnvironmentSatisfaction)

# Repeat all modifications for CS2 (Training set) into 2 Test datasets
# Convert categorical data to factors

TestAtt$Over18 <- NULL
TestAtt$StandardHours <- NULL
TestAtt$EmployeeCount <- NULL
TestAtt$EmployeeNumber <- NULL

TestSal$Over18 <- NULL
TestSal$StandardHours <- NULL
TestSal$EmployeeCount <- NULL
TestSal$EmployeeNumber <- NULL

TestAtt$JobInvolvement <- as.factor(TestAtt$JobInvolvement)
TestAtt$JobLevel <- as.factor(TestAtt$JobLevel)
TestAtt$JobSatisfaction <- as.factor(TestAtt$JobSatisfaction)
TestAtt$PerformanceRating <- as.factor(TestAtt$PerformanceRating)
TestAtt$RelationshipSatisfaction <- as.factor(TestAtt$RelationshipSatisfaction)
TestAtt$StockOptionLevel <- as.factor(TestAtt$StockOptionLevel)
TestAtt$TrainingTimesLastYear <- as.factor(TestAtt$TrainingTimesLastYear)
TestAtt$WorkLifeBalance <- as.factor(TestAtt$WorkLifeBalance)
TestAtt$Education <- as.factor(TestAtt$Education)
TestAtt$EnvironmentSatisfaction <- as.factor(TestAtt$EnvironmentSatisfaction)

TestSal$JobInvolvement <- as.factor(TestSal$JobInvolvement)
TestSal$JobLevel <- as.factor(TestSal$JobLevel)
TestSal$JobSatisfaction <- as.factor(TestSal$JobSatisfaction)
TestSal$PerformanceRating <- as.factor(TestSal$PerformanceRating)
TestSal$RelationshipSatisfaction <- as.factor(TestSal$RelationshipSatisfaction)
TestSal$StockOptionLevel <- as.factor(TestSal$StockOptionLevel)
TestSal$TrainingTimesLastYear <- as.factor(TestSal$TrainingTimesLastYear)
TestSal$WorkLifeBalance <- as.factor(TestSal$WorkLifeBalance)
TestSal$Education <- as.factor(TestSal$Education)
TestSal$EnvironmentSatisfaction <- as.factor(TestSal$EnvironmentSatisfaction)

# Check Data Structure

str(CS2)
summary(CS2)
str(TestAtt)
str(TestSal)
```


# Plot the Data
## Check Adherence to MLR Assumptions: Normality, Equal Variance, Independence, Outliers and Leverage.

```{r}
plot(CS2$Attrition, CS2$MonthlyIncome)
```

### Divide Training Dataset into smaller groups before plotting

####  People-Oriented Variables:  Age, Education, EducationField, Gender, MaritalStatus, RelationshipSatisfaction, TrainingTimesLastYear

####  Income-oriented variables:  DailyRate, HourlyRate, JobInvolvement, JobLevel, JobSatisfaction, MonthlyIncome, MonthlyRate, OverTime, PercentSalaryHike, StockOptionLevel

####  Organizational variables:  Department, JobRole, NumCompaniesWorked, PerformanceRating, TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager, Attrition

####  Work-Life Balance variables:  Attrition,BusinessTravel, DistanceFromHome, EnvironmentSatisfaction, WorkLifeBalance

```{r}

Peopledf <- CS2 %>% select(c(Age, Education, EducationField, Gender, MaritalStatus, RelationshipSatisfaction, TrainingTimesLastYear))
Peopledf %>% ggpairs()

Incomedf <- CS2 %>% select(c(DailyRate, HourlyRate, JobInvolvement, JobLevel, JobSatisfaction, MonthlyIncome, 
                             MonthlyRate, OverTime, PercentSalaryHike, StockOptionLevel))
Incomedf %>% ggpairs()

Orgdf <- CS2 %>% select(c(Department, JobRole, NumCompaniesWorked, PerformanceRating, TotalWorkingYears,
                          YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager, Attrition))
Orgdf %>% ggpairs()

WLdf <- CS2 %>% select(c(Attrition,BusinessTravel, DistanceFromHome, EnvironmentSatisfaction, WorkLifeBalance))
WLdf %>% ggpairs()


# Numeric data distribution - histograms from skim
skim(CS2)  # from library(skimr)
```

### Analyze Collinearity among numeric factors - Are any numeric relationships close to 1 or -1?

```{r}

correlator  <-  function(df){
  df %>%
    keep(is.numeric) %>%
    tidyr::drop_na() %>%
    cor %>%
    corrplot("upper", addCoef.col = "white", number.digits = 2,
             number.cex = 0.5, method="square",
             order="hclust", title="Correlation of Numeric Variables",
             tl.srt=45, tl.cex = 0.8)
}

correlator(CS2)

# Variable selection
# General - Are any variables closely correllated with each other?
# Years* values have approx 70% correllation so be aware if both are in the resulting model(s)
# 1.  Which variables are most significantly tied to Attrition? (classification)
# 2.  Which variables will best model the relationship to MonthlyIncome?  (muliple linear regression)
#


# Question 1 Salary Analysis

# Used caret package to run forward, backward and stepwise variable selection for linear regression
# Tuned number of variables with tuneGrid and nvmax
# use nvmax = 3 to accommodate Frito-Lay requirement of top 3 factors


# Salary:  Forward regression
# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.modelF <- caret::train(MonthlyIncome ~., data = CS2,
                            method = "leapForward", 
                            tuneGrid = data.frame(nvmax = 68),
                            trControl = train.control
)
step.modelF$results
step.modelF$bestTune
summary(step.modelF$finalModel)

# Salary:  Backward Regression
# Set seed for reproducibility
set.seed(456)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)

step.modelB <- caret::train(MonthlyIncome ~., data = CS2,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 61),
                    trControl = train.control
)
step.modelB$results
step.modelB$bestTune
summary(step.modelB$finalModel)

# Salary:  Stepwise regression
# Set seed for reproducibility
set.seed(789)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.modelS <- caret::train(MonthlyIncome ~., data = CS2,
                            method = "leapSeq", 
                            tuneGrid = data.frame(nvmax = 56),
                            trControl = train.control
)
step.modelS$results
step.modelS$bestTune
summary(step.modelS$finalModel)
coef(step.modelS$finalModel, 56)

# Salary:  Custom Model with reduced number of factors

SalaryVars <- c("MonthlyIncome","Age", "Gender", "JobLevel", "JobRole", "BusinessTravel","StockOptionLevel", "TotalWorkingYears")
SalaryCS2 <- CS2[SalaryVars]

# Plot the data
SalaryCS2 %>% ggpairs() 

# Set seed for reproducibility
set.seed(333)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)

train.model <- caret::train(MonthlyIncome ~ ., data=SalaryCS2, 
                     method = "leapSeq", 
                     tuneGrid = data.frame(nvmax = 3),
                     trControl = train.control
)
train.model$results
train.model$bestTune
train.model$finalModel
coef(train.model$finalModel,3)
summary(train.model$finalModel)

# Salary:  Use lm and reduced dataframe - best R^2
```
## Develop Linear Models for Salary Analysis
### Custom variable selection model with lm linear regression was best performing

## Run Naive Bayes classification method with two different models.
### Reduced variables, tied to Attrition were best overall model.

# Conclusion
##  EDA is time consuming and it is difficult to see relationships with categorical variables.
##  Salary was strongly correlated to job role, job level and a few other variables.
##  Attrition was tied to number of companies worked, years since last promotion, travel vs. salary.
##  Recommend deeper analysis on job role and job level as well as pay for travel time.
##  GenderMale was 90 points higher than reference for the model.

# The End.  Thank you!

```{r}
SalaryCustomLM <- lm(MonthlyIncome ~ ., data=SalaryCS2 )
summary(SalaryCustomLM)



# Question 2:  Predict Attrition (Yes or No) based on Classification models
#

# Begin Classification Work - Naive Bayes on CS2  = Full dataset minus variables that are unique to each person.

# Create Train and Test datsets from CS2 master
# split data into 70% training and 30% test data sets
# From Unit 7 Dr. Sadler's code

################ This model works but the specificity is a bit low
set.seed(2)
trainInd = sample(seq(1,dim(CS2)[1],1),round(.7*dim(CS2)[1]))
CS2Train = CS2[trainInd,]
CS2Test = CS2[-trainInd,]

set.seed(2)
trainInd = sample(seq(1,dim(CS2)[1],1),round(.7*dim(CS2)[1]))
CS2Train = CS2[trainInd,]
CS2Test = CS2[-trainInd,]

model = naiveBayes(Attrition ~ .,data = CS2Train)
table(predict(model,CS2Test),CS2Test$Attrition)
CM = confusionMatrix(table(predict(model,CS2Test),CS2Test$Attrition))
CM


# Custom variable selection for Attrition - All performance metrics  > 60%

AttritionVars <- c("Attrition","Age", "Department", "MonthlyIncome","MaritalStatus","BusinessTravel","DistanceFromHome","EnvironmentSatisfaction","JobInvolvement","JobLevel","OverTime","PerformanceRating","StockOptionLevel","TrainingTimesLastYear","WorkLifeBalance")

AttritionCS2 <- CS2[AttritionVars]
str(AttritionCS2)

set.seed(297)
trainInd = sample(seq(1,dim(AttritionCS2)[1],1),round(.7*dim(CS2)[1]))
AttritionCS2Train = CS2[trainInd,]
AttritionCS2Test = CS2[-trainInd,]

model2 = naiveBayes(Attrition ~ .,data = AttritionCS2Train)
table(predict(model2,AttritionCS2Test),AttritionCS2Test$Attrition)
CM2 = confusionMatrix(table(predict(model2,AttritionCS2Test),AttritionCS2Test$Attrition))
CM2


##########################################################
# Apply models to test datasets for NoAttrition and NoSalary
# Output to .csv files and put in Github
# File Output code provided by Malcolm Carlson, advised by his Professor
#

# Attrition Output

AttModel <- naiveBayes(Attrition ~., data = AttritionCS2)

AttPred <- predict(AttModel,TestAtt)
myout=cbind.data.frame(TestAtt$ID,AttPred)
colnames(myout) <- c("ID","Attrition")
write.csv(myout, file ="C:/SMU MS Data Science Files/SMU Downloads/6306 DDS/Project 2 DDSAnalytics/Case2PredictionsLazarescouAttrition.csv", row.names = FALSE)

# Salary output

SalaryModel <- lm(MonthlyIncome ~ ., data=SalaryCS2)

SalaryPred <- predict(SalaryModel,TestSal)
myout=cbind.data.frame(TestSal$ID,SalaryPred)
colnames(myout) <- c("ID","MonthlyIncome")
write.csv(myout, file ="C:/SMU MS Data Science Files/SMU Downloads/6306 DDS/Project 2 DDSAnalytics/Case2PredictionsLazarescouSalary.csv", row.names = FALSE)



```


